# Programming With R and Python

## R Programming (Part 1)

**Note: Ctrl + Shift + K**

### Problem 1

For $a=5, b=15$ and $c =10$, find

-   $$\frac{a+b}{c}$$

-   $$\frac{-b+\sqrt{b^2 - 4ac}}{2a}$$

-   $$ac+|c-b|$$

-   $$(2a-c)^2$$

### Solution

```{r}
a <- 5
b <- 15
c <- 10

# (i) 
(a+b)/c
# (ii)
(-b+sqrt(b^2-4*a*c)) / (2*a)
# (iii)
a*c + abs(c-b)
# (iv)
(2*a - c)^2
```

------------------------------------------------------------------------

### Problem 2

i)  Create a data vector in R for 20,13,14,25,16,20,10,9,23,35,70,12,15,13,10. Manipulate the data vector by removing 70.

ii) Create a data vector in R which is 1 to 15

iii) Create a data vector in R which is 1 to 24 by 2

iv) Create a following data matrix in R $$  \begin{bmatrix} 
      1 & 2 & 3 & 4 \\
      5 & 6 & 7 & 8 \\
      9 & 10 & 11 & 12 
      \end{bmatrix} $$ Also find (a) row sum, (b) column sum and (c) the element (2,3)

### Solution:

```{r}
# (i)
x <- c(20,13,14,25,16,20,10,9,23,35,70,12,15,13,10)
x <- x[-11] # all the elements without position 11
print(x)

# (ii) Data vector (1 to 15)
x2 <- 1:15
print(x2)

# (iii) Data vector (1 to 24 by 2)
x3 <- seq(1, 24, 2)
print(x3)

# (iv) Matrix
A <- matrix(1:12, nrow=3, byrow = T)
print(A)
# (a) row sums
rowSums(A)
# (b) column sums
colSums(A)
# element (2,3)
A[2,3]
```

------------------------------------------------------------------------

### Problem 3

i)  Create the following data matrix in R $$ \begin{bmatrix}
      1 & 2 & 3\\
      5 & 6 & 7\\
      9 & 10 & 11
    \end{bmatrix}$$

    a)  Transpose of A
    b)  Find the inverse of A, and
    c)  Eigen values of A

ii) Consider another matrix $$\begin{bmatrix} 
      7 & 4 & 1\\
      5 & 6 & 7\\
      2 & 7 & 3
    \end{bmatrix}$$

    a)  Find A+B
    b)  Find A x B

### Solution

```{r}
# (i)
A <- matrix(1:12, 3, 4, T)
dimnames(A) <- list(paste0("c", 1:3), paste0("r", 1:4))
A
A <- A[,-4]
print(A)

# (a) Transpose of A
t(A)
# (b)	Find the inverse of A
# solve(A) => Error
# (c)	Eigen values of A
eigen(A)

# (ii)
B <- matrix(c(7,5,2,4,6,7,1,7,3), 3, 3)
dimnames(B) <- list(paste0("c", 1:3), paste0("r", 1:3))
# (a) A+B, A-B
A+B
A-B
# (b) A*B
A %*% B 
```

------------------------------------------------------------------------

### Problem 4

#### Consider the simple fictitious data-set that describes patient responses to two drugs at five dosage levels as given below.

| Dosage | Response to Drug A | Response to Drug B |
|:-------|:-------------------|:-------------------|
| 20     | 16                 | 15                 |
| 30     | 20                 | 18                 |
| 40     | 27                 | 25                 |
| 45     | 40                 | 31                 |
| 60     | 60                 | 40                 |

**Plot response of the two drugs against dosage.** *[add title, legend, text of axes, and set a reference line at 30]*

### Solution

```{r}
dose  <- c(20, 30, 40, 45, 60)
drugA <- c(16, 20, 27, 40, 60)
drugB <- c(15, 18, 25, 31, 40)
color <- c("red", "blue", "purple", "navy", "gray")
.pch <- c(15, 17)
.lty <- c(1, 2)
plot(dose, drugA, type="b", pch=.pch[1], lty=.lty[1], col=color[1], ylim=c(0, 60), xlab="", ylab="")
title(main = "Drug A vs. Drug B", col.main=color[3],
      xlab="Drug Dosage", ylab="Drug Response", col.lab=color[4])
lines(dose, drugB, type="b", pch=.pch[2], lty=.lty[2], col=color[2])
abline(h=c(15, 30, 45), lwd = 1.5, lty = 3, col=color[5])

legend("topleft", inset=0.05, title="Drug Type", c("A","B"), lty=.lty, pch=.pch, col=color[1:2])
```


### Problem 5

 i) Use Newton’s method for finding root: $f(x) = x^3 + 2x^2 - 7$
 ii) Solve the quadratic equation $x^2 + 2*x - 7 =0$ using if-else(), and repeat().
 iii) Write a function to compute n! and hence find 23!
 iv) Compute co-efficient of covariance of the data set $c(5, 6,2 ,8,5,9,2,4,8,3,1,9)$
 
 
#### Solution
 
*(i) $f(x) = x^3 + 2x^2 - 7$*
 ```{r}
x <- 1
f <- x^3 + 2*x^2 - 7
tolerance <- 0.000001

while(abs(f) > tolerance ) {
  f.prime <- 3*x^2 + 4*x
  x <- x - f/f.prime
  f <- x^3 + 2*x^2 - 7
  print(x)
}
 ```
 
 *(ii) Solve the quadratic equation $x^2 + 2x - 7 = 0$ using if-else(), and repeat().*
 
 ```{r}
 
 ```